Lightweight Ruby Framework for Improving Embedded Software Efficiency

### 2016-06-09 CPSNA2016/paper.tex
update CPSNA2016

% \title{Efficient Devlopment Framework applied Scripting Language for Embedded Systems}
\title{Efficient Development Framework with mruby\\for Embedded Systems}
\author{\IEEEauthorblockN{Takuro Yamamoto}\IEEEauthorblockA{Graduate School of Engineering Science,\\Osaka University}
\and\IEEEauthorblockN{Hiroshi Oyama}\IEEEauthorblockA{OKUMA Corporation}
\and\IEEEauthorblockN{Takuya Azumi}\IEEEauthorblockA{Graduate School of Engineering Science,\\Osaka University}
}

\begin{document}
\maketitle
\begin{abstract}
Recently, the productivity of embedded systems applied to Cyber Physical Systems has become problematic due to their increasing complexity and scale.
To improve the productivity, the mruby on TOPPERS embedded component system (TECS) framework, which employs scripting language (i.e., mruby) and supports component-based development, has been proposed.
This paper proposes an extended framework of mruby on TECS that improve development efficiency more than the current version.
In the proposed framework, a Bluetooth loader which receives an mruby bytecode, and a RiteVM scheduler which simplifies multitasking are implemented.
% In the current mruby on TECS framework, mruby programs must be compiled and linked every time they are modified, because mruby bytecode are incorporated in the platform.
% Moreover, while the framework supports multiple virtual machines, developers must be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs concurrently or in parallel.
% To improve development efficiency, we propose a Bluetooth loader for mruby bytecode as an extension of mruby on TECS.
% The loader executes two mruby bytecodes, i.e., the mruby application bytecode and mruby library bytecode.
% The mruby application bytecode is modified frequently and sent from a host to a target device by the developers, while the mruby library bytecode is modified infrequently and preserved beforehand on a storage/ROM device during the first compilation.
% In addition, multiple mruby programs can run cooperatively in the proposed framework.
% Compared to mruby on TECS, the RiteVM scheduler simplifies multitasking.
% Synchronization of initializing multiple tasks is also implemented using an Eventflag. 
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}

% \keywords{embedded software; scripting language; component-based development}
\begin{IEEEkeywords}
embedded software; scripting language; component-based development
\end{IEEEkeywords}

\section{Introduction}
Cyber Physical Systems (CPS) is the important keyword for enriching our lives.
Embedded systems e.g., Internet of Things applications, are applied to the CPS and expected to demonstrate high quality and high performance.
This requirement has led to an increase in their complexity and scale; moreover, these systems need to have low production costs and short development cycles.

@@

To address the above limitation, ``mruby on TECS,'' a component-based framework for running script programs, has been proposed \cite{par:mrubyonTECS}.
This framework integrates two technologies, i.e., mruby, which is a lightweight implementation of Ruby for embedded systems \cite{par:mruby}, and TECS, which is a component-based framework for embedded systems \cite{par:TECS}.
% mruby on TECS supports to effectively run mruby script language on embedded systems.
% mruby on TECS also makes execution time 100 times faster than that of mruby.

%This paper proposes two additional features of mruby on TECS, Bluetooth loader for mruby bytecode and RiteVM scheduler.
Even though execution times of mruby on TECS are 100 times faster than those of mruby, it is not particularly efficient, at present, and imposes a heavy burden on developers.
mruby on TECS only supports a storage/ROM device for loading mruby programs.
Consequently, if mruby programs are modified, an SD card must be inserted and removed repeatedly or ROM must be rewritten; moreover, developers need to restart an RTOS on the target device.
In addition, although mruby on TECS can support multi-VM, executing multiple tasks requires the developers to call the OS function.

@@

{\bf Contributions}: The proposed framework provides the following contributions:
\begin{enumerate}
    \vspace{-2mm}
\item {\bf Improved software development efficiency.}
Developers do not need to rewrite a storage/ROM device and restart an RTOS.
% The Bluetooth loader supports continuous loading, which reduces Bluetooth set-up time (i.e., pairing).
\item {\bf Execution of multiple mruby programs concurrently or in parallel.}
Developers can implement multiple tasks without RTOS knowledge because the RiteVM scheduler switches tasks cyclically.
\item {\bf Synchronized execution of multiple RiteVM tasks.}
The proposed framework synchronizes multiple RiteVM tasks (i.e., mruby applications).
\item {\bf Benefits of CBD:}
The paper focuses on the benefits of CBD and provides specific examples.
    \vspace{-2mm}
\end{enumerate}

@@

\subsection{Bluetooth Loader for mruby Bytecode}
\label{sec:Bluetooth loader for mruby bytecode}
This section describes the proposed additional functionality, i.e., the Bluetooth loader.\footnote{
The Bluetooth loader is intended to improve development efficiency; therefore, software developers should use it during the development phase.
Note that the complete software should be compiled and linked on the storage/ROM device beforehand.
}
In the current system, the platform including mruby bytecodes is saved on a storage/ROM device.
Developers must rewrite the storage/ROM device every time the programs are modified.
In addition, the RTOS on the target device needs to be restarted.
The repetition hinders development efficiency.
The Bluetooth loader for mruby bytecode decreases developer burden because developers only have to connect the storage/ROM device and start the RTOS once.

@@

We evaluated to demonstrate that a Bluetooth loader can improve the efficiency of software development, that the proposed multitask processing executes effectively compared to singletasking or co-routine, and that the initiation of mruby applications are synchronized.
In addition, we focused on benefits of CBD.
We implemented the proposed system on a LEGO MINDSTORMS EV3 \cite{par:EV3} (300MHz ARM9-based Sitara AM1808 system-on-a-chip) compiled with gcc 4.9.3 -O2 and mruby version 1.2.0.

{\bf Improving Development Efficiency by Bluetooth Loader}:
The Bluetooth loader can reduce the development time.
With the proposed framework, developers do not need to rewrite a storage/ROM device because only the bytecode should be transferred.
For example, with the existing system, when mruby programs are modified, developers must remove an SD card from the target device, connect the host PC, compile/link the platform, and reinsert the SD card in the target device in an experimental environment such as LEGO MINDSTORMS EV3.
In addition, the proposed framework eliminates the need to restart the target RTOS.

@@

{\bf RiteVM Scheduler}:
A comparison of the application execution time with singletasking, co-routine, and multitasking is shown in Figure \ref{fig:comparison_s_c_m}.
We decided a program with 100,000 loops as an mruby application for the evaluation of execution time.
Here, the singletask program looped 100,000 times and the multitask and co-routine programs looped 50,000 times for each task.

\begin{figure}[t]
    \centering
    \includegraphics[height=4.0cm,width=8.6cm,clip]{figure/comparison_s_c_m.eps}
    \vspace{-3mm}
\caption{Comparison of application execution time}
    \vspace{-3mm}
\label{fig:comparison_s_c_m}
\end{figure}
\begin{figure}[t]
    \centering
    \includegraphics[height=4.0cm,width=8.6cm,clip]{figure/comparison_msec.eps}
    \vspace{-3mm}
\caption{Comparison of overhead for each cyclic period of RiteVM scheduler}
    \vspace{-3mm}
\label{fig:comparison_msec}
\end{figure}
\subsubsection{Execution Time of Singletasking, Co-routine, and Proposed Multitasking}
Figure \ref{fig:comparison_s_c_m} shows that the proposed design is superior to co-routine in terms of execution time.
Moreover, developers can utilize the scheduler practically because the RiteVM scheduler overhead is approximately 5\%.
The scheduler interrupts and switches tasks, which causes this overhead.
Although the overhead is larger as the number of RiteVMs increases, it also spread the difference between multitasking and co-routine.
Therefore, the scheduler enables developers multitasking as the more efficient approach than co-routine.
Here, the periodic time of the scheduler is 1 msec.
% It takes approximately 3 $\mu$sec on average to switch tasks once.

@@

 \begin{figure}[t]
    \centering
    \includegraphics[height=4.0cm,width=8.6cm,clip]{figure/eval_synchronization.eps}
    \vspace{-3mm}
\caption{Synchronization of multiple RiteVM tasks}
    \vspace{-3mm}
\label{fig:eval_synchronization}
\end{figure}

@@

{\bf Synchronization of Multiple RiteVM Tasks}:
To execute multiple mruby applications, a synchronization mechanism for RiteVM tasks is implemented in the proposed framework.
We measured the time from the execution of the first RiteVM task to the execution of the last RiteVM task.
It was confirmed that the time was within (periodic time)$\times$(number of RiteVM tasks $-$ 1).
Figure \ref{fig:eval_synchronization} shows the results for two, three, and four RiteVM tasks.
% Figure \ref{fig:eval_synchronization} shows the case of four RiteVM tasks.
% The periodic time is one msec, and the number of RiteVM is four.
The periodic time is 1 msec, and the number of RiteVMs is two, three, and four.
As shown in Figure \ref{fig:eval_synchronization}, the time is within 1 msec, 2 msec, and 3 msec respectively, which indicates successful synchronization of multiple RiteVM tasks.

{\bf Benefits of Component-Based Development}:
In the proposed framework, RiteVMs, the RiteVM scheduler, and Eventflags are implemented as TECS components.
Developers can add or remove the functionalities easily by modifying the CDL file.
Moreover, CBD decreases code size and improves productivity and maintainability.

@@

\begin{table}[t]
    \centering
    % \vspace{-2mm}
    \caption{C and CDL file code for the number of RiteVMs}
    \vspace{-2mm}
    %\scriptsize
    {\tabcolsep=0.3cm
    \begin{tabular}{l||c|c|c}
                & (A)       & (B)     & Diff  \\ \hline
        C (Total)      & 8$\times$$\alpha$$+$134  & 130     & 8$\times$$\alpha+$4\\
        C (Modification)   & 10$\times\alpha$$-$2 & 0   &  10$\times\alpha$$-$2 \\
        CDL    & 18$\times$$\alpha$$+$25   & 18$\times$$\alpha$$+$25 & 0     \\
        \multicolumn{3}{l}{{\small $\alpha$} : {\scriptsize the number of RiteVM}}
    \end{tabular}
    }
          \vspace{-3mm}
    \label{tab:codesize}
\end{table}

@@

{\bf mruby:} mruby, a lightweight implementation of the Ruby language, has been proposed for embedded systems.
mruby programs can run on a RiteVM, which is the VM for mruby and reads the mruby bytecode.
Note that the RiteVM only supports a single thread.
In addition, mruby supports co-routines but does not support multitasking for RTOSs.

@@

\section{Conclusion}
\label{sec:Conclusion}
This paper presents an extended framework of mruby on TECS, i.e., the proposed framework implements a Bluetooth loader for mruby bytecode and a RiteVM scheduler.
The Bluetooth loader improves software development efficiency by eliminating the need to rewrite a storage/ROM device and restart an RTOS.
The proposed framework can be applied to various embedded systems because the loader can use both Bluetooth and a wired serial connection.
%Thus, the proposed framework can be utilized on the devices without Bluetooth.
The RiteVM scheduler simplifies multitasking compared to the current mruby on TECS.
Our experimental results for the Bluetooth loader and RiteVM scheduler show their advantages.
The Bluetooth loader can improve software development efficiency on mruby on TECS, and the RiteVM scheduler is more effective than co-routines for multitasking.
Moreover, synchronization of multiple RiteVM tasks is implemented in the proposed framework.

@@

\newpage
\appendices
\section{mruby}
mruby is a light-weight implementation of the Ruby programming language complying to part of the ISO standard.
Ruby is an object-oriented scripting language with classes and methods, exceptions, and garbage collection functions.
It is easy to use and read due to its simple grammar and Ruby requires fewer lines of code than C.
Ruby improves the productivity of software development due to its simple grammar and object-oriented functions.

>>>>

% \title{Efficient Devlopment Framework applied Scripting Language for Embedded Systems}
\title{Lightweight Ruby Framework for Improving Embedded Software Efficiency}
% \title{Efficient Development Framework with mruby\\for Embedded Systems}
\author{\IEEEauthorblockN{Takuro Yamamoto}\IEEEauthorblockA{Graduate School of Engineering Science,\\Osaka University}
\and\IEEEauthorblockN{Hiroshi Oyama}\IEEEauthorblockA{OKUMA Corporation}
\and\IEEEauthorblockN{Takuya Azumi}\IEEEauthorblockA{Graduate School of Engineering Science,\\Osaka University}
}

\begin{document}
\maketitle
\begin{abstract}
Recently, the productivity of embedded systems applied to Cyber Physical Systems has become problematic due to their increasing complexity and scale.
To improve the productivity, the mruby on TOPPERS embedded component system (TECS) framework, which employs scripting language (i.e., a lightweight Ruby) and supports component-based development, has been proposed.
This paper proposes an extended framework of mruby on TECS that improve development efficiency more than the current version.
In the proposed framework, a Bluetooth loader which receives an mruby bytecode, and a RiteVM scheduler which simplifies multitasking are implemented.
% In the current mruby on TECS framework, mruby programs must be compiled and linked every time they are modified, because mruby bytecode are incorporated in the platform.
% Moreover, while the framework supports multiple virtual machines, developers must be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs concurrently or in parallel.
% To improve development efficiency, we propose a Bluetooth loader for mruby bytecode as an extension of mruby on TECS.
% The loader executes two mruby bytecodes, i.e., the mruby application bytecode and mruby library bytecode.
% The mruby application bytecode is modified frequently and sent from a host to a target device by the developers, while the mruby library bytecode is modified infrequently and preserved beforehand on a storage/ROM device during the first compilation.
% In addition, multiple mruby programs can run cooperatively in the proposed framework.
% Compared to mruby on TECS, the RiteVM scheduler simplifies multitasking.
% Synchronization of initializing multiple tasks is also implemented using an Eventflag. 
% Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}

% \keywords{embedded software; scripting language; component-based development}
\begin{IEEEkeywords}
scripting language; cyber physical systems; embedded software; component-based development
\end{IEEEkeywords}

\section{Introduction}
Cyber Physical Systems (CPS) are the important keyword for enriching our lives.
Embedded systems e.g., Internet of Things applications, are applied to the CPS and expected to demonstrate high quality and high performance.
This requirement has led to an increase in their complexity and scale; moreover, these systems need to have low production costs and short development cycles.

@@

To address the above limitation, ``mruby on TECS,'' a component-based framework for running script programs, has been proposed \cite{par:mrubyonTECS}.
This framework integrates two technologies, i.e., mruby, which is a lightweight implementation of Ruby for embedded systems \cite{par:mruby}, and TECS, which is a component-based framework for embedded systems \cite{par:TECS}.
% mruby on TECS supports to effectively run mruby script language on embedded systems.
% mruby on TECS also makes execution time 100 times faster than that of mruby.
%This paper proposes two additional features of mruby on TECS, Bluetooth loader for mruby bytecode and RiteVM scheduler.
Execution times of mruby on TECS are 100 times faster than those of mruby, which are the same as those of C language.
Moreover, Ruby is often utilized in network programming, therefore the framework can be applicable for the software involved network such as CPS applications.

However, it is not particularly efficient, at present, and imposes a heavy burden on developers.
mruby on TECS only supports a storage/ROM device for loading mruby programs.
Consequently, if mruby programs are modified, an SD card must be inserted and removed repeatedly or ROM must be rewritten; moreover, developers need to restart an RTOS on the target device.
In addition, although mruby on TECS can support multi-VM, executing multiple tasks requires the developers to call the OS function.

@@

{\bf Contributions}: The proposed framework provides the following contributions:
\begin{enumerate}
    % \vspace{-2mm}
\item {\bf Improved software development efficiency.}
Developers do not need to rewrite a storage/ROM device and restart an RTOS.
% The Bluetooth loader supports continuous loading, which reduces Bluetooth set-up time (i.e., pairing).
\item {\bf Execution of multiple mruby programs concurrently or in parallel.}
Developers can implement multiple tasks without RTOS knowledge because the RiteVM scheduler switches tasks cyclically.
\item {\bf Synchronized execution of multiple RiteVM tasks.}
The proposed framework synchronizes multiple RiteVM tasks (i.e., mruby applications).
\item {\bf Benefits of CBD.}
The paper focuses on the benefits of CBD and provides specific examples (a Bluetooth loader, a sheduler, and Eventflags).
    % \vspace{-2mm}
\end{enumerate}

@@

\subsection{Bluetooth Loader for mruby Bytecode}
\label{sec:Bluetooth loader for mruby bytecode}
This section describes the proposed additional functionality, i.e., the Bluetooth loader.\footnote{
The Bluetooth loader is intended to improve development efficiency; therefore, software developers should use it during the development phase.
Note that the completed software production should be compiled and linked on the storage/ROM device beforehand.
}
In the current system, the platform including mruby bytecodes is saved on a storage/ROM device.
Developers must rewrite the storage/ROM device every time the programs are modified.
In addition, the RTOS on the target device needs to be restarted.
The repetition hinders development efficiency.
The Bluetooth loader for mruby bytecode decreases developer burden because developers only have to connect the storage/ROM device and start the RTOS once. 
An existing Bluetooth-based loader such as TOPPERS/EV3RT \cite{par:EV3} has supported a load of object files.
The proposed loader novelly supports a load of bytecodes.

@@

We evaluated to demonstrate that a Bluetooth loader can improve the efficiency of software development, that the proposed multitask processing executes effectively compared to existing multitask processing or co-routine, and that the initiation of mruby applications are synchronized.
In addition, we focused on benefits of CBD.
We implemented the proposed system on a LEGO MINDSTORMS EV3 \cite{par:EV3} (300MHz ARM9-based Sitara AM1808 system-on-a-chip) compiled with gcc 4.9.3 -O2 and mruby version 1.2.0.

{\bf Improving Development Efficiency by Bluetooth Loader.}
The Bluetooth loader can reduce the development time.
With the proposed framework, developers do not need to rewrite a storage/ROM device because only the bytecode should be transferred.
For example, with the existing system, when mruby programs are modified, developers must remove an SD card from the target device, connect the host PC, compile/link the platform, and reinsert the SD card in the target device in an experimental environment such as LEGO MINDSTORMS EV3.
In addition, the proposed framework eliminates the need to restart the target RTOS.

@@

{\bf RiteVM Scheduler.}
A comparison of the application execution time with singletask, proposed, existing multitask, and co-routine is shown in Figure \ref{fig:comparison_s_c_m}.
We applied a program with 100,000 loops as an mruby application for the evaluation of execution time so that the overhead comes out.
Here, in the multitask and co-routine, we prepared two tasks and the programs looped 50,000 times for each task.

\begin{figure}[t]
    \centering
    \includegraphics[height=3.0cm,width=7.0cm,clip]{figure/comparison_s_c_m.eps}
    \vspace{-3mm}
\caption{Comparison of application execution time}
    \vspace{-3mm}
\label{fig:comparison_s_c_m}
\end{figure}
\begin{figure}[t]
    \centering
    \includegraphics[height=3.0cm,width=7.0cm,clip]{figure/comparison_msec.eps}
    \vspace{-3mm}
\caption{Comparison of overhead for each cyclic period of RiteVM scheduler}
    \vspace{-3mm}
\label{fig:comparison_msec}
\end{figure}
\subsubsection{Execution Time of Singletask, Proposed Multitask, Existing Multitask, and Co-routine}
Figure \ref{fig:comparison_s_c_m} shows that the proposed design is superior to multitask using delay method and co-routine in terms of execution time.
Moreover, developers can utilize the scheduler practically because the RiteVM scheduler overhead is approximately 5\%.
The scheduler interrupts and switches tasks, which causes this overhead.
Although the overhead is larger as the number of RiteVMs increases, it also spread the difference between multitask and co-routine.
Therefore, the scheduler enables developers multitask as the more efficient approach than the existing multitasking and co-routine.
Here, the periodic time of the scheduler is 1 msec.
% It takes approximately 3 $\mu$sec on average to switch tasks once.

@@

 \begin{figure}[t]
    \centering
    \includegraphics[height=3.0cm,width=7.0cm,clip]{figure/eval_synchronization.eps}
    \vspace{-3mm}
\caption{Synchronization of multiple RiteVM tasks}
    \vspace{-3mm}
\label{fig:eval_synchronization}
\end{figure}

@@

{\bf Synchronization of Multiple RiteVM Tasks.}
To execute multiple mruby applications, a synchronization mechanism for RiteVM tasks is implemented in the proposed framework.
We measured the time from the execution of the first RiteVM task to the execution of the last RiteVM task.
It was confirmed that the time was within (periodic time)$\times$(number of RiteVM tasks $-$ 1).
Figure \ref{fig:eval_synchronization} shows the results for two, three, and four RiteVM tasks.
% Figure \ref{fig:eval_synchronization} shows the case of four RiteVM tasks.
% The periodic time is one msec, and the number of RiteVM is four.
The periodic time is 1 msec, and the number of RiteVMs is two, three, and four.
As shown in Figure \ref{fig:eval_synchronization}, the time is within 1 msec, 2 msec, and 3 msec respectively, which indicates successful synchronization of multiple RiteVM tasks.

{\bf Benefits of Component-Based Development.}
In the proposed framework, RiteVMs, the RiteVM scheduler, and Eventflags are implemented as TECS components.
Developers can add or remove the functionalities easily by modifying the CDL file.
Moreover, CBD decreases code size and improves productivity and maintainability.

@@

\begin{table}[t]
    \centering
    % \vspace{-2mm}
    \caption{C and CDL file code for the number of RiteVMs}
    \vspace{-2mm}
    %\scriptsize
    {\tabcolsep=0.3cm
    \begin{tabular}{l||c|c|c}
                & (A)       & (B)     & Diff  \\ \hline
        C (Total)      & 8$\times$$\alpha$$+$134  & 130     & 8$\times$$\alpha+$4\\
        C (Modification)   & 10$\times\alpha$$-$2 & 0   &  10$\times\alpha$$-$2 \\
        CDL    & 18$\times$$\alpha$$+$25   & 18$\times$$\alpha$$+$25 & 0     \\
        \multicolumn{3}{l}{{\small $\alpha$} : {\scriptsize the number of RiteVM}}
        \vspace{-3mm}
    \end{tabular}
    }
          \vspace{-3mm}
    \label{tab:codesize}
\end{table}

@@

{\bf mruby:} mruby, a lightweight Ruby, has been proposed for embedded systems.
mruby programs can run on a RiteVM which reads the mruby bytecode.
Note that the RiteVM only supports a single thread.
In addition, mruby supports co-routines but does not support multitasking for RTOSs.

@@

\section{Conclusion}
\label{sec:Conclusion}
This paper presents an extended framework of mruby on TECS, i.e., the proposed framework implements a Bluetooth loader for mruby bytecode and a RiteVM scheduler.
The Bluetooth loader improves software development efficiency by eliminating the need to rewrite a storage/ROM device and restart an RTOS.
The proposed framework can be applied to various embedded systems because the loader can use both Bluetooth and a wired serial connection.
%Thus, the proposed framework can be utilized on the devices without Bluetooth.
The RiteVM scheduler simplifies multitasking compared to the current mruby on TECS.
Our experimental results for the Bluetooth loader and RiteVM scheduler show their advantages.
The Bluetooth loader can improve software development efficiency on mruby on TECS, and the RiteVM scheduler is more effective than the existing technique and co-routines for multitasking.
Moreover, synchronization of multiple RiteVM tasks is implemented in the proposed framework.

@@

\clearpage
\appendices
\section{mruby}
mruby is a light-weight implementation of the Ruby programming language complying to part of the ISO standard.
Ruby is an object-oriented scripting language with classes and methods, exceptions, and garbage collection functions.
It is easy to use and read due to its simple grammar and Ruby requires fewer lines of code than C.
Ruby improves the productivity of software development due to its simple grammar and object-oriented functions.
